Report on Late Winter2013 Production: Image Differencing

### 2013-03-18 W13report.tex
Spell checking

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We use variations on the
configuration where we use one Gaussian, with the above width, and
three Gaussians, the central of which has this width, and the others a
factor of $\beta = 2.0$ smaller and larger.  We modify each of the
Gaussians by a set of Laguerre polynomials of a given order.  The
smallest Gaussian is modified to the highest order, with the others
modified by order//2.  Figure~\ref{basis} provides an example basis
set with this configuration.  The total number of shapes in the basis
are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription to basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 2.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convovled the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  Hoewver, it is not clear
  that devonvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.
}.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 5                      & 29  \\
         & False     & 10                     & 81  \\
v6866602 & True      & 5                      & 30  \\
         & False     & 9                      & 35  \\
v6866603 & True      & 2                      & 9  \\
         & False     & 1                      & 23  \\
\end{tabular}
\caption{Total number of {\tt DiaSources} from all 9 sensors in raft
  2,2, for the configurations that yielded the lowest number of false
  positives.  Matches are determiend using a 2'' match radius with the
  input reference catalog.  Objects with the {\tt EDGE} flag are
  ignored. \label{tab-bestfp}}
\end{table}

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Configs} \\
\hline
Visit    & Prefilter & nGauss & degGauss & spatialOrder \\
\hline
v6866601 & True      & 3      & 6        & 5 \\
         & False     & 3      & 2        & 4 \\
v6866602 & True      & 3      & 6        & 5 \\
         & False     & 3      & 5        & 5 \\
v6866603 & True      & 3      & 6        & 4 \\
         & False     & 3      & 6        & 4 \\
\end{tabular}
\caption{Configuations that led to the best results listed in
  Table~\ref{tab-bestfp}.  All post-production runs include
  perturbations about these configurations. \label{tab-bestconfig}}
\end{table}

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/falsePos.eps} \\
\caption{``Heat--map'' detailing the total numbers of false positives
  as a function of two main degrees of freedom: the maximum order of
  basis Laguerre polynomials along the y--axis, and the spatial order
  of the Chebyshev expansion of the spatial model along the x--axis.
  The {\it left} column shows these results for prefiltering, while
  the {\it right} column shows this for the post--filtered analysis.
  The best results listed in Table~\ref{tab-bestfp}, with the
  associated configureations from Table~\ref{tab-bestconfig},
  correspond to the minimum in each of the subpanels.
}
\label{fp_heatmap}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/deconv2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/size2.eps} \\
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/order2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/shape2.eps} \\
\caption{Four example panels showing failure modes of the image
  subtraction software.  Each panel itself shows a 2x2 pair of
  difference images; on the {\it left} subpanel is the residual from
  the LOCAL fitting, and on the {\it right} from the SPATIAL fitting.
  Starting in the upper left and going clockwise, the first panel
  shows the residuals when devonvolving the template image; the
  ringing is characteristic of the devonvolution process.  Next, a set
  of difference images where the kernel dimensions (kernelSize) are
  too small to match the sources, leaving square kernel--sized
  residuals around each object.  Next, a set of images where the
  kernel shape (sigGauss) is inappropriate to match the sources; note
  that the LOCAL residuals show circular residuals, indicating the
  basis set itself is at fault (as opposed to the spatial model).
  Finally, in the lower--left, a set of images that demonstrate how
  the residuals degrade when the spatial kernel model is at fault.
  Note that the LOCAL residuals appear to be white noise, while the
  SPATIAL residuals show a clear dipole signature.  For this reason, a
  comparison beteen the LOCAL and SPATIAL residuals is a useful
  diagnostic of the spatial model.  }
\label{fig_galleryBad}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/normal2.eps} \\
\includegraphics[width=0.9\textwidth]{figures/preconv2.eps} \\
\end{center}
\caption{Pairs of LOCAL,SPATIAL difference images in the case of
  successful image subtraction, similar to
  Figure~\ref{fig_galleryBad}.  The {\it top} set of data are from
  postfiltered analysis, while the {\it lower} set are from
  prefiltered data.}
\label{fig_galleryGood}
\end{figure}

>>>>

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We use variations on the
configuration where we use one Gaussian, with the above width, and
three Gaussians, the central of which has this width, and the others a
factor of $\beta = 2.0$ smaller and larger.  We modify each of the
Gaussians by a set of Laguerre polynomials of a given order.  The
smallest Gaussian is modified to the highest order, with the others
modified by order//2.  Figure~\ref{basis} provides an example basis
set with this configuration.  The total number of shapes in the basis
are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription to basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 2.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.
}.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{clcc}
\hline
\multicolumn{4}{|c|}{Best Results} \\
\hline
Visit    & Prefilter & False Positive Matches & False Positive Orphans \\
\hline
v6866601 & True      & 5                      & 29  \\
         & False     & 10                     & 81  \\
v6866602 & True      & 5                      & 30  \\
         & False     & 9                      & 35  \\
v6866603 & True      & 2                      & 9  \\
         & False     & 1                      & 23  \\
\end{tabular}
\caption{Total number of {\tt DiaSources} from all 9 sensors in raft
  2,2, for the configurations that yielded the lowest number of false
  positives.  Matches are determined using a 2'' match radius with the
  input reference catalog.  Objects with the {\tt EDGE} flag are
  ignored. \label{tab-bestfp}}
\end{table}

\begin{table}
\centering
\begin{tabular}{clccc}
\hline
\multicolumn{5}{|c|}{Best Configs} \\
\hline
Visit    & Prefilter & nGauss & degGauss & spatialOrder \\
\hline
v6866601 & True      & 3      & 6        & 5 \\
         & False     & 3      & 2        & 4 \\
v6866602 & True      & 3      & 6        & 5 \\
         & False     & 3      & 5        & 5 \\
v6866603 & True      & 3      & 6        & 4 \\
         & False     & 3      & 6        & 4 \\
\end{tabular}
\caption{Configurations that led to the best results listed in
  Table~\ref{tab-bestfp}.  All post-production runs include
  perturbations about these configurations. \label{tab-bestconfig}}
\end{table}

@@

\begin{figure}
\includegraphics[width=1.0\textwidth]{figures/falsePos.eps} \\
\caption{``Heat--map'' detailing the total numbers of false positives
  as a function of two main degrees of freedom: the maximum order of
  basis Laguerre polynomials along the y--axis, and the spatial order
  of the Chebyshev expansion of the spatial model along the x--axis.
  The {\it left} column shows these results for prefiltering, while
  the {\it right} column shows this for the post--filtered analysis.
  The best results listed in Table~\ref{tab-bestfp}, with the
  associated configurations from Table~\ref{tab-bestconfig},
  correspond to the minimum in each of the subpanels.
}
\label{fp_heatmap}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/deconv2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/size2.eps} \\
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/order2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/shape2.eps} \\
\caption{Four example panels showing failure modes of the image
  subtraction software.  Each panel itself shows a 2x2 pair of
  difference images; on the {\it left} subpanel is the residual from
  the LOCAL fitting, and on the {\it right} from the SPATIAL fitting.
  Starting in the upper left and going clockwise, the first panel
  shows the residuals when deconvolving the template image; the
  ringing is characteristic of the deconvolution process.  Next, a set
  of difference images where the kernel dimensions (kernelSize) are
  too small to match the sources, leaving square kernel--sized
  residuals around each object.  Next, a set of images where the
  kernel shape (sigGauss) is inappropriate to match the sources; note
  that the LOCAL residuals show circular residuals, indicating the
  basis set itself is at fault (as opposed to the spatial model).
  Finally, in the lower--left, a set of images that demonstrate how
  the residuals degrade when the spatial kernel model is at fault.
  Note that the LOCAL residuals appear to be white noise, while the
  SPATIAL residuals show a clear dipole signature.  For this reason, a
  comparison between the LOCAL and SPATIAL residuals is a useful
  diagnostic of the spatial model.  }
\label{fig_galleryBad}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/normal2.eps} \\
\includegraphics[width=0.9\textwidth]{figures/preconv2.eps} \\
\end{center}
\caption{Pairs of LOCAL,SPATIAL difference images in the case of
  successful image subtraction, similar to
  Figure~\ref{fig_galleryBad}.  The {\it top} set of data are from
  postfiltered analysis, while the {\it lower} set are from
  prefiltered data.  Note that the noise is smoother in the
  prefiltered data due to correlation with the Psf. }
\label{fig_galleryGood}
\end{figure}
